Data Management for LSST Special Programs

### 2023-10-17 sec_sci.tex
Update sec_sci.tex
partial updates

\section{Maximizing Science from Special Programs}\label{sec:sci}

High-level recommendations regarding RDM implementations to meet the 
requirements related to Special Programs summarized in Section~\ref{sec:req}.

\subsection{Labeling for Queryability}\label{ssec:sci_labels}

As described in Section~\ref{sec:req}, there is a requirement that metadata for 
observations associated with Special Programs is stored and is sufficient for 
triggering real-time data processing recipes.

Additionally, the Special Program science goals require that users are able to 
query and retrieve processed image and catalog data by specifiying the Special Program.

Labelling is also essential for provenance and mini-survey 
characterizability (estimate selection biases), and for Special Programs 
processing by RDM or by users.

\textbf{All visits should have a region label \textit{and} a program 
label, and these labels should be propagated to all processed images and catalog data.}

\textbf{Region Labels -- } based on Right Acension and Declination.
Region labels should include, e.g., WFD low-dust, Galactic Plane, Galactic 
Bulge, SCP, NES, or Virgo; DDF fieldname; mini-survey region.
Region labels should be propagated to all processed images (visit, 
difference, and deep coadds) and all cataloged sources and objects.

\textbf{Program Labels -- } 
Should be based on survey cadeance or strategy.
Program labels might include, e.g., WFD, DDF, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Program labels should be propagated to all single-visit processed and 
difference images, and all cataloged sources.

@@

The existence and propagation of these labels is not a requirement.
This document suggests that these labels be the adopted implementation for 
the metadata requirements (see Section~\ref{ssec:req_meta}).

\subsection{Processing and Data Products}

As described in Section~\ref{sec:req}, there is a requirement that 
Rubin Observatory produce unique, separate, and joinable data products 
whenever this is possible with the original or reconfigured versions of the 
LSST Science Pipelines (and no new algorithmic development or significant 
computational resources are needed).

From a science perspective, however, in some cases where it is \emph{possible} 
to produce such separate data products it is not \emph{necessary}; these
cases are discussed below.

\textbf{Case A (possible but not necessary due to inclusion in WFD data products) -- } 
For some science goals, although it would be possible to produce separate 
data products for a Special Program with the LSST Science Pipelines, this 
is not necessary because the visits can be processed with, and included in, 
the WFD data products (Section~\ref{ssec:proc_wfd}).
These cases are usually related to time domain astronomy, difference image
analysis (DIA), prompt processing, and alert production.
Users would simply incorporate the region and program labels in queries to restrict 
their analysis to only a certain Special Program (or only WFD) where the 
science goal requires it.

@@

\subsection{Computational Resources for User Processing}

@@

\subsection{Rubin Science Platform Capabilities}

>>>>

\section{Recommended Implementations to Enable Science from Special Programs}\label{sec:sci}

Recommendations regarding implementations by Rubin Data Management to meet the 
requirements related to Special Programs summarized in Section~\ref{sec:req},
which will enable science with data from Special Programs.

\subsection{Region and Program Labels}\label{ssec:sci_labels}

\textbf{All visits should have a region label \textit{and} a program 
label, and these labels should be propagated to processed images and catalog data.}

\textbf{Region Labels -- } based on Right Acension and Declination.
Region labels should include, e.g., WFD low-dust, Galactic Plane, Galactic 
Bulge, SCP, NES, or Virgo; DDF fieldname; mini-survey region name.
Region labels should be propagated to processed images (visit, 
difference, and deep coadds), catalog sources, and catalog static-sky objects.

\textbf{Program Labels -- } 
Based on survey cadeance or strategy.
Program labels might include, e.g., WFD, DDF, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Program labels should be propagated to single-visit processed and 
difference images and catalog sources.

@@

The existence and propagation of these labels is not itself a requirement.
It is recommended for implementation for three reasons.

\begin{enumerate}

\item To meet the requirement that metadata for observations associated 
with Special Programs is stored, and is sufficient for triggering 
real-time data processing recipes (Section~\ref{ssec:req_meta}).

\item To enable users to query and retrieve processed image and catalog
data associated with a specific Special Program, and meet the science goals
of that Special Program, when regular processing as been applied 
(e.g., Prompt Processing, Section~\ref{ssec:sci_pproc}).

\item To enable provenance when Special Programs data in included in regular
processing and the WFD program data products (e.g., if used to improve the
all-sky coadd).

\end{enumerate}

\subsection{Special Processing}\label{ssec:sci_sproc}

\textbf{Special Processing should be done by Rubin Data Management to 
produce unique and separate (but joinable) data products
for Special Programs when it is both possible and necessary.}

\textbf{Possible -- } When original or reconfigured versions of the LSST
Science Pipelines can be used, and no new algorithmic or software development,
or significant additional computational resources, are needed.

\textbf{Necessary -- } When the science goals for a Special Program cannot
be met by only including the data in regular processing (e.g., Prompt processing).

\begin{enumerate}

\item To meet the requirement that Rubin Observatory produce
unique, separate, and joinable data products whenever this is possible 
with the original or reconfigured versions of the LSST Science Pipelines
(Section~\ref{sec:req}).

\item To enable science with Special Programs by all users, not just those
with the time and effort to process the data, and to reduce computational
load (and potential redundancy) in User-Generated Processing.

\end{enumerate}

To further illustrate this recommendation, a few examples of Special Processing
cases are provided below.

\begin{itemize}

\item \textbf{Possible and necessary:}

\item \textbf{Possible but not necessary:}
a time-domain mini-survey that uses standard visits \emph{could}
have separate difference-image analysis object and source catalogs
generated, but this is not necessary as the science goals for the
mini-survey can be met by processing its data with regular Prompt
Processing (Section~\ref{ssec:sci_pproc}), and ensuring the
data is properly labeled (Section~\ref{ssec:sci_labels}).

\item \textbf{Necessary but not possible:}

\end{itemize}

RESTART EDITING PROCESS HERE

@@

\subsection{Prompt Processing}\label{ssec:sci_pproc}

\textbf{Special Processing should produce unique and separate data products
for Special Programs when it is both possible and necessary.}

\subsection{Computational Resources for User-Generated Processing}\label{ssec:sci_comp}

@@

\subsection{Rubin Science Platform Capabilities}\label{ssec:sci_rsp}
